#2.2 THe very basics

Before we get started with the motivating dataset, we need to cover the very basics of R.

##2.2.1 Objects
Suppose a high school student asks us for help solving several quadratic equations of the form $ax^2+bx+c=0$.The quadratic formula gives us the solutions:
$$
\frac{-b-\sqrt{b^2-4ac}}{2a} and \frac{-b+\sqrt{b^2-4ac}}{2a}
$$
which of course change depending on the values of $a$, $b$, and  $c$. One advantage of programming languages is that we can define variables and write expressions with these variables, similar to how we do so in math, but obtain a numeric solution. We will write out general code for the quadratic equation below, but if we are asked to solve $x^2+X-1=0$, then we define:
```{r}
a <- 1
b <- 1
c <- -1
```
which stores the values for later use. We use <- to assign values to the variables.

We can also assign values using = instead of <-, but we recommend against using = to avoid confusion.

Copy and paste the code above into your console to define the three variables. Note that R does not print anything when we make this assignment. This means the objects were defined successfully. Had you made a mistake, you would have received an error message.

To see the value stored in a variable, we simply ask R to evaluate a and it shows the stored value:
```{r}
a
```


